---
title: "STA 567 HW5"
author: "Lina Lee"
date: "10/25/2019"
output: word_document
---
```{r include=FALSE}
library(tidyverse)
library(MASS)
library(splines)
library(caret)
```

```{r}
setwd("C:/Users/linal/Desktop/Miami2019/STA567/Homework/Homework5")
powerplant <- read.csv("powerplant.csv")
```

# Validation choice
For all the models, I used 5-fold cross validation to evaluate models’ predictive performance.


# Polynomial regression
```{r}
# The first order polynomial
set.seed(123)
mlr_mod1 <- train(PE ~ V, 
                   data=powerplant,
                   method="lm",
                   trControl=trainControl(method="cv",number = 5),
                   preProcess = c("center", "scale"))

mean(mlr_mod1$resample$RMSE)

# The second order polynomial
set.seed(123)
mlr_mod2 <- train(PE ~ V + I(V^2), 
                   data=powerplant,
                   method="lm",
                   trControl=trainControl(method="cv",number = 5),
                   preProcess = c("center", "scale"))

mean(mlr_mod2$resample$RMSE)

# The third order polynomial
set.seed(123)
mlr_mod3<- train(PE ~ V + I(V^2) + I(V^3), 
                   data=powerplant,
                   method="lm",
                   trControl=trainControl(method="cv",number = 5),
                   preProcess = c("center", "scale"))

mlr_mod3$results$RMSE

# The fourth order polynomial
set.seed(123)
mlr_mod4<- train(PE ~ V + I(V^2) + I(V^3) +I(V^4), 
                   data=powerplant,
                   method="lm",
                   trControl=trainControl(method="cv",number = 5),
                   preProcess = c("center", "scale"))

mlr_mod4$results$RMSE

```


# Knot choice.
 I refer to the scatter plot to make four sets of three knots.I chose three sets of knots around 45,55,and 70 since the pattern of the data spread changes at those points points on the scatterplot. 

```{r}
#create 4 sets of knots
v_knots1 <- c(min(powerplant$V),40,65,70,max(powerplant$V))
v_knots2 <- c(min(powerplant$V),45,55,70,max(powerplant$V))
v_knots3 <- c(min(powerplant$V),45,58,65,max(powerplant$V))
v_knots4 <- c(min(powerplant$V),50,60,75,max(powerplant$V))

# create a knots list inluding four sets of knots.
knot_list<-list(v_knots1,v_knots2,v_knots3,v_knots4)

#create a function to make a bin for each knots set
knots<-function(v_knots){
bin_v <- cut(powerplant$V, breaks=v_knots,
                          right=FALSE, include.lowest=TRUE)
return(bin_v)
}

#apply the function to create bin to knots list
bin_list<-lapply(knot_list,function(x) knots(x))

# transform bin list into dataframe
bin_df<-as.data.frame(bin_list)

# Change variable name for the each bin
names(bin_df)<-c("bin1", "bin2", "bin3", "bin4")

# combine bin columns with original data
data_bins<-cbind(powerplant,bin_df)
```


# piecewise regression using 3 knots
```{r}
# Try the first knots
set.seed(123)
  pwr_mod <- train(PE ~ bin1*V +bin1*I(V^2), 
                   #use the combined data including bin
                   data=data_bins,
                   method="lm",
                   trControl=trainControl(method="cv",number = 5),
                   preProcess = c("center", "scale"))
 
  pwr_mod$results$RMSE

```

```{r}
# Try the second knots
set.seed(123)
  pwr_mod2 <- train(PE ~ bin2*V +bin2*I(V^2), 
                   data=data_bins,
                   method="lm",
                   trControl=trainControl(method="cv",number = 5),
                   preProcess = c("center", "scale"))

  pwr_mod2$results$RMSE

```

```{r}
# Try the third knots
set.seed(123)
  pwr_mod3 <- train(PE ~ bin3*V +bin3*I(V^2), 
                   data=data_bins,
                   method="lm",
                   trControl=trainControl(method="cv",number = 5),
                   preProcess = c("center", "scale"))

  pwr_mod3$results$RMSE

```


```{r}
# Try the fourth knots
set.seed(123)
  pwr_mod4 <- train(PE ~ bin4*V +bin4*I(V^2), 
                   data=data_bins,
                   method="lm",
                   trControl=trainControl(method="cv",number = 5),
                   preProcess = c("center", "scale"))

  pwr_mod4$results$RMSE

```


# Cubic regression splines 
```{r warning=FALSE}
# create a bs matrix
bs<-bs(powerplant$V,knots=v_knots1,degree=3)
?bs
bs_mat<-as.data.frame(bs)
names(bs_mat)<-c("x","x2","x3","x_kn1","x_kn2","x_kn3","x_kn4","x_kn5")
# combine the bs matrix with original powerplant data
powerplant2<-cbind(powerplant,bs_mat)

# Try the first knots
set.seed(123)
power_spline1 <- train(PE~x+x2+x3+x_kn1+x_kn2+x_kn3+x_kn4+x_kn5,
                 data=powerplant2,
                 method="lm",
                 trControl=trainControl(method="cv",number = 5),
                 preProcess = c("center", "scale"))

power_spline1$results$RMSE
```
```{r  warning=FALSE}
# create a bs matrix
bs<-bs(powerplant$V,knots=v_knots2,degree=3)
bs_mat<-as.data.frame(bs)
names(bs_mat)<-c("x","x2","x3","x_kn1","x_kn2","x_kn3","x_kn4","x_kn5")
# combine the bs matrix with original powerplant data
powerplant2<-cbind(powerplant,bs_mat)

# Try the second knots
set.seed(123)
power_spline2 <- train(PE~x+x2+x3+x_kn1+x_kn2+x_kn3+x_kn4+x_kn5,
                 data=powerplant2,
                 method="lm",
                 trControl=trainControl(method="cv",number = 5),
                 preProcess = c("center", "scale"))

power_spline2$results$RMSE
```

```{r warning=FALSE}
# create a bs matrix
bs<-bs(powerplant$V,knots=v_knots3,degree=3)
bs_mat<-as.data.frame(bs)
names(bs_mat)<-c("x","x2","x3","x_kn1","x_kn2","x_kn3","x_kn4","x_kn5")
# combine the bs matrix with original powerplant data
powerplant2<-cbind(powerplant,bs_mat)

# Try the third knots
set.seed(123)
power_spline3 <- train(PE~x+x2+x3+x_kn1+x_kn2+x_kn3+x_kn4+x_kn5,
                 data=powerplant2,
                 method="lm",
                 trControl=trainControl(method="cv",number = 5),
                 preProcess = c("center", "scale"))

power_spline3$results$RMSE
```

```{r warning=FALSE}
# create a bs matrix
bs<-bs(powerplant$V,knots=v_knots4,degree=3)
bs_mat<-as.data.frame(bs)
names(bs_mat)<-c("x","x2","x3","x_kn1","x_kn2","x_kn3","x_kn4","x_kn5")
# combine the bs matrix with original powerplant data
powerplant2<-cbind(powerplant,bs_mat)

# Try the fourth knots
set.seed(123)
power_spline4 <- train(PE~x+x2+x3+x_kn1+x_kn2+x_kn3+x_kn4+x_kn5,
                 data=powerplant2,
                 method="lm",
                 trControl=trainControl(method="cv",number = 5),
                 preProcess = c("center", "scale"))

power_spline4$results$RMSE
```

# Smoothing splines
```{r warning=FALSE}
# The first df trial: df=seq(0,1000,by=20)
set.seed(123)
smooth_spline1<- train(PE ~ V, 
                  method="gamSpline",
                  data=powerplant, 
                  trControl=trainControl(method="cv",number = 5),
                  tuneGrid=data.frame(df=seq(0,1000,by=20)))

smooth_spline1$finalModel$tuneValue
```

```{r warning= FALSE}
# The second df trial: df=seq(120,160,by=5)
set.seed(123)
smooth_spline2<- train(PE ~ V, 
                  method="gamSpline",
                  data=powerplant, 
                  trControl=trainControl(method="cv",number = 5),
                  tuneGrid=data.frame(df=seq(120,160,by=5)))

smooth_spline2$finalModel$tuneValue
```

```{r warning=FALSE}
# The last df trial: df=seq(120,130,by=1)
set.seed(123)
smooth_spline3<- train(PE ~ V, 
                  method="gamSpline",
                  data=powerplant, 
                  trControl=trainControl(method="cv",number = 5),
                  tuneGrid=data.frame(df=seq(120,130,by=1)))

smooth_spline3$finalModel$tuneValue
mean(smooth_spline3$resample$RMSE)
```
First, I tried sequence from 0 to 1000 by 20 for degrees of freedom to investigate best tune. The result gave me the best df of 140. Next, I applied the sequence from 120 to 160 by 5 for degrees of freedom. The result says that df 125 is the best tune.I investigate further by using the sequence from 120 to 130 by 1 for the degrees of freedom. From the result, the degrees of freedom 124 is the best.

# QQ: span is almost 0...?

# LOESS using a standard weight
```{r}
# The first tune grid
tune_grid <- expand.grid(span = seq(0.1, 0.9, by=0.2), degree = 1)

set.seed(123)
loess_mod<- train(PE ~ V, 
                      method="gamLoess",
                      data=powerplant, 
                      trControl=trainControl(method="cv",number = 5),
                      tuneGrid=tune_grid)

loess_mod$finalModel$tuneValue
```

```{r}
# The second tune grid
tune_grid2 <- expand.grid(span = seq(0.05, 0.2, by=0.1), degree = 1)

set.seed(123)
loess_mod2<- train(PE ~ V, 
                      method="gamLoess",
                      data=powerplant, 
                      trControl=trainControl(method="cv",number = 5),
                      tuneGrid=tune_grid2)

loess_mod2$finalModel$tuneValue
```

```{r}
# The last tune grid 
tune_grid3 <- expand.grid(span = seq(0.01, 0.1, by=0.005), degree = 1)

set.seed(123)
loess_mod3<- train(PE ~ V, 
                      method="gamLoess",
                      data=powerplant, 
                      trControl=trainControl(method="cv",number = 5),
                      tuneGrid=tune_grid2)

loess_mod3$finalModel$tuneValue
mean(loess_mod3$resample$RMSE)
```
### span and polynomial degree choice
First, I tried sequence from 0.1 to 0.9 by 0.2, and degree 1. The degree 2 broke all the operations, so I only applied degree 1. The result said that span 0.1 is the best tune.since span 0.1 is edge of the sequence, I investigate further below the span 0.1. secondly, I tried the sequence from 0.05 to 0.2 by 0.1. The result said that span 0.05 is the best. Finally, I applied the sequence from 0.01 to 0.1 by 0.005. The result said that span 0.05 is the best tune. 0.05 is not a edge in the sequence, I didn't investigate furthur, and I concluded that span 0.05 is the best tune.





