---
title:  "STA567 Homework7"
author: "Lina Lee"
date: "11/9/2019"
output:
  word_document: default
  html_document: default
---


```{r,include=FALSE}
# Load the library
library(pdfCluster)
library(caret)
# Load data for accessing
data("oliveoil")
```

# Using 10-fold cross validation, fit and tune a support vector classifier for predicting the region of each oil sample.
```{r}
tc <- trainControl("repeatedcv", number=10, repeats=10)
```
# SVC linear kernal
```{r ,cache=TRUE}
svc_linear_oliveoil <- train(region ~ ., 
                            data=oliveoil,
                            method="svmLinear",
                            trControl=tc,
                            tuneGrid=data.frame(C=c(1,10,50,100)))

svc_linear_oliveoil$results[which.max(svc_linear_oliveoil$results$Accuracy),]
```
## Further investigate the tunring parameter
```{r ,cache=TRUE}
svc_linear_oliveoil2 <- train(region ~ ., 
                            data=oliveoil,
                            method="svmLinear",
                            trControl=tc,
                            tuneGrid=data.frame(C=c(0.5,1,3,5)))

svc_linear_oliveoil2$results[which.max(svc_linear_oliveoil2$results$Accuracy),]
# misclassification rate
1-max(svc_linear_oliveoil2$results$Accuracy)
```
```{r}
# fit the final model to make a confusion matrix below.
svc_linear_final <- train(region ~ ., 
                            data=oliveoil,
                            method="svmLinear",
                            tuneGrid=data.frame(C=1))

```


# SVC polynomial kernal
```{r ,cache=TRUE}
svc_poly_oliveoil <- train(region ~ ., 
                          data=oliveoil,
                          method="svmPoly",
                          trControl=tc,
                          tuneGrid=expand.grid(C=c(.1,1,10,100),
                                               degree=2:4, scale=1))
# The best Tune and corresponding Accuracy
svc_poly_oliveoil$results[which.max(svc_poly_oliveoil$results$Accuracy),]
```

### Further investigate the tunring parameter
```{r ,cache=TRUE}

svc_poly_oliveoil2 <- train(region ~ ., 
                          data=oliveoil,
                          method="svmPoly",
                          trControl=tc,
                          tuneGrid=expand.grid(C=c(0.01,0.05,0.08,1),
                                               degree=2:4, scale=1))

svc_poly_oliveoil2$results[which.max(svc_poly_oliveoil2$results$Accuracy),]

# The misclassification rate
1-max(svc_poly_oliveoil$results$Accuracy)

```
```{r}
# 0.05	2
svc_poly_final <- train(region ~ ., 
                          data=oliveoil,
                          method="svmPoly",
                          trControl=tc,
                          tuneGrid=expand.grid(C=0.05,
                                               degree=2, scale=1))


```

# SVC radial kernal
```{r ,cache=TRUE}
svc_rad_oliveoil <- train(region ~ ., 
                         data=oliveoil,
                         method="svmRadial",
                         trControl=tc,
                         tuneGrid=expand.grid(C=c(1,2,4,8,16,30),
                         sigma=c(0.01,0.1,0.16,0.64,0.99)))
# The best Tune and corresponding Accuracy
svc_rad_oliveoil$results[which.max(svc_rad_oliveoil$results$Accuracy),]
```

## Further investigate the tunring parameter
```{r,cache=TRUE}
svc_rad_oliveoil2 <- train(region ~ ., 
                         data=oliveoil,
                         method="svmRadial",
                         trControl=tc,
                         tuneGrid=expand.grid(C=c(1,2,4,8,16),
                         sigma=c(0.05,0.08,0.1,0.16,0.2,0.4)))
# The best Tune and corresponding Accuracy
svc_rad_oliveoil2$results[which.max(svc_rad_oliveoil2$results$Accuracy),]

# The misclassification rate
1-max(svc_rad_oliveoil$results$Accuracy)
```

```{r}
svc_rad_final<- train(region ~ ., 
                         data=oliveoil,
                         method="svmRadial",
                         trControl=tc,
                         tuneGrid=expand.grid(C=2,
                         sigma=0.1))

```

# Report any tuning parameters chosen, as well as the overall misclassification rate. 

SVM classifier| Model Parameters      |Misclassification rate 
--------------| ----------------------|------
   Linear     |      C=1              |0.04110
 Polynomial   |C=0.05,degree=2,scale=1|0.04132
Radial Kernels|   C=9,sigma=0.08      |0.03787

Among three classifier, Radial Kernels has lowest Misclassification rate. Therefore, the SVM with Radial Kernels perfromed best among the three classifiers. 

# From the confusion matrix, are there some regions where the classifier performs particularly well?
```{r}
library(kableExtra)
x <- subset(oliveoil, select=-region)
y<- oliveoil$region
pred_linear <- predict(svc_linear_final,x)

linear_mat<-table(pred_linear,y)
kable(linear_mat)
```

In the regions of Apulia.north,Sicily Sardinia.inland, Sardinia.coast, Liguria.west, Umbria, all the predicted region match with the actual region.
the SVC Linear model perfroms well in the regions. 

```{r}
pred_poly<- predict(svc_poly_final,x)
poly_mat<-table(pred_poly,y)
kable(poly_mat)
```

In the regions of Sicily, Sardinia.inland, Sardinia.coast, Liguria.east, Liguria.west, Umbria, all the predicted region match with the actual regions.
the SVC polynomial model perfroms well in the regions. 

```{r}
pred_rad<- predict(svc_rad_final,x)
rad_mat<-table(pred_rad,y)
kable(rad_mat)
```

In the regions of Apulia.north, Sicily, Sardinia.inland, Sardinia.coast, Liguria.east, Liguria.west, Umbria, all the predicted region match with the actual regions.the SVC radial kernel model perfroms well in thoese regions. 