---
title: "STA567 HW4"
author: "Lina Lee"
date: "10/4/2019"
output: word_document
---

```{r,results='hide'}

setwd("C:\\Users\\linal\\Desktop\\Miami2019\\STA567\\Homework\\Homework4")
load(file="Decathlons.Rdata")
head(london)

# 100m run, long jump, shotput, high jump  and 400m run
# x100m long shotput high x400m
# x110m x1500m
# London-trained
```

```{r,results='hide'}
library(tidyverse)
library(caret)
```

# Remove missing values
```{r}
london <- london %>%
  select(x110m, x1500m, x100m, long, shotput, high, x400m) %>%
  filter(!is.na(x110m)) %>%
  filter(!is.na(x1500m)) 
```


# (1) Backward Stepwise Regression from AIC
## Model for 110m hurdles
```{r}
# Backward Stepwise
mod1 <- lm(x110m ~ .-x1500m,data=london)
stepBackward <- step(mod1)
stepBackward

# Cross Validation 
set.seed(12345)
mlr1 <- train(x110m ~ x100m + long + shotput + high, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))


# RMSE for 5th-fold cross validation
min(mlr1$results$RMSE)  
```

## Model for 1500m run
```{r}
# Backward Stepwise Regression from AIC
mod2 <- lm(x1500m ~ .-x110m,data=london)
stepBackward <- step(mod2)
stepBackward

# Cross Validation
set.seed(12345)
mlr2 <- train(x1500m ~ x100m + long + shotput + x400m, 
              data=london,
              method="lm",
              trControl=trainControl(method="cv",number = 5),
              preProcess = c("center", "scale"))

# RMSE for 5th-fold cross validation
min(mlr2$results$RMSE) 
```

# (2) Lasso regression 
## Model for 110m hurdles
```{r}
# Set seed for reproducibility
set.seed(12345)
# Train the model
lasso_mod1<-train(x110m ~ .-x1500m , 
                 data=london,
                 method="lasso",
                 # Set up repeated k-fold cross-validation
                 trControl=trainControl(method="cv",number=5),
                 preProcess = c("center","scale"),
                 tuneLength=20)

lasso_mod1$bestTune
mean(lasso_mod1$resample$RMSE)
```

## Model for 1500m run
```{r}
### x1500m
set.seed(12345)
lasso_mod2<-train(x1500m ~ .-x110m , 
                 data=london,
                 method="lasso",
                 trControl=trainControl(method="cv",number=5),
                 preProcess = c("center", "scale"),
                 tuneLength=20)

lasso_mod2$bestTune
mean(lasso_mod2$resample$RMSE)
```

# (3) Rigde regression
## Model for 110m hurdles
```{r}
set.seed(12345)
ridge_mod1 <- train(x110m ~ .-x1500m , 
                   data=london,
                   method="ridge",
                   trControl=trainControl(method="cv",number = 5),
                   preProcess = c("center", "scale"),
                   tuneLength=20)

ridge_mod1$bestTune
min(ridge_mod1$results$RMSE)
```


## Model for 1500m run
```{r}
set.seed(12345)
ridge_mod2 <- train(x1500m ~ .-x110m , 
                    data=london,
                    method="ridge",
                    trControl=trainControl(method="cv",number = 5),
                    preProcess = c("center", "scale"),
                    tuneLength=20)


# depending on the model, criteria to choose bestTUne is different? not RMSE?? 
ridge_mod2$bestTune
min(ridge_mod2$results$RMSE)
```



# (4) elastic net 
## Model for 110m hurdles
```{r}
set.seed(12345)
enet_mod1 <- train(x110m ~ .-x1500m , 
                  data=london,
                  method="enet",
                  trControl=trainControl(method="cv",number = 5),
                  preProcess = c("center", "scale"),
                  tuneLength=20)

enet_mod1$bestTune
min(enet_mod1$results$RMSE)
```


## Model for 1500m run
```{r}
set.seed(12345)
enet_mod2 <- train(x1500m ~ .-x110m , 
                   data=london,
                   method="enet",
                   trControl=trainControl(method="cv",number = 5),
                   preProcess = c("center", "scale"),
                   tuneLength=20)

enet_mod2$bestTune
min(enet_mod2$results$RMSE)
```


# (5) Principal Component Regression
## Model for 110m hurdles
```{r}
set.seed(12345)
pcr_mod1 <- train(x110m ~ .-x1500m, 
                 data=london, 
                 method="pcr",
                 preProcess=c("center","scale"),
                 trControl = trainControl(method="cv",number = 5),
                 tuneGrid = data.frame(ncomp=1:6))


pcr_mod1
pcr_mod1$bestTune
min(pcr_mod1$results$RMSE)
```


## Model for 1500m run
```{r}
set.seed(12345)
pcr_mod2 <- train(x1500m ~ .-x110m, 
                 data=london, 
                 method="pcr",
                 preProcess=c("center","scale"),
                 trControl = trainControl(method="cv",number = 5),
                 tuneGrid = data.frame(ncomp=1:6))

pcr_mod2$bestTune
min(pcr_mod2$results$RMSE)
```




# (6) Partial Least Squares Regression
## Model for 110m hurdles
```{r}
set.seed(12345)
plsr_mod1 <- train(x110m ~ .-x1500m, 
                  data=london, 
                  method="pls",
                  preProcess=c("center","scale"),
                  trControl = trainControl(method="cv"),
                  tuneGrid = data.frame(ncomp=1:6))


plsr_mod1$bestTune
min(plsr_mod1$results$RMSE)
```

## Model for 1500m run
```{r}
set.seed(12345)
plsr_mod1 <- train(x1500m ~ .-x110m, 
                  data=london, 
                  method="pls",
                  preProcess=c("center","scale"),
                  trControl = trainControl(method="cv"),
                  tuneGrid = data.frame(ncomp=1:6))


plsr_mod1$bestTune
min(plsr_mod1$results$RMSE)
```


# (7) Backward Stepwise Regression from RMSE 
### full model
```{r}
set.seed(12345)
mod <- train(x110m ~ x100m+ long+ shotput+ high+ x400m, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
min(mod$results$RMSE)
```

### STEP1 
### Drop one variable from the full model
```{r}
set.seed(12345)
mod1 <- train(x110m ~ x100m+ long+ shotput+ high, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE1<-min(mod1$results$RMSE)  

set.seed(12345)
mod2 <- train(x110m ~ x100m+ long+ shotput+ x400m, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE2<-min(mod2$results$RMSE)  

set.seed(12345)
mod3 <- train(x110m ~ x100m+ long+ high+ x400m, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE3<-min(mod3$results$RMSE)  

set.seed(12345)
mod4 <- train(x110m ~ x100m+ shotput+ high+ x400m, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))          
            
RMSE4<-min(mod4$results$RMSE)  

set.seed(12345)
mod5 <- train(x110m ~ long+ shotput+ high+ x400m, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE5<-min(mod5$results$RMSE) 

RMSE_list<-c(RMSE1,RMSE2,RMSE3,RMSE4,RMSE5)
RMSE_list
min(RMSE_list)
```

*mod 1 has least RMSE 0.4322. Now our improved model is lm(x110m ~ x100m+ long+ shotput+ high)*


### STEP2
### Drop one variable from the improved model from STEP1.
```{r}
set.seed(12345)
mod2_1 <- train(x110m ~ x100m+ long+ shotput, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE2_1<-min(mod2_1 $results$RMSE)  

set.seed(12345)
mod2_2 <- train(x110m ~ x100m+ long+ high, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE2_2<-min(mod2_2$results$RMSE)  

set.seed(12345)
mod2_3 <- train(x110m ~ x100m+ shotput+ high, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE2_3<-min(mod2_3$results$RMSE)  

set.seed(12345)
mod2_4 <- train(x110m ~ long+ shotput+ high, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE2_4<-min(mod2_4$results$RMSE)  

RMSE2_list<-c(RMSE2_1,RMSE2_2,RMSE2_3,RMSE2_4)
RMSE2_list
min(RMSE2_list)
```
The second model lm(x110m ~ x100m+ long+ high) in step2 has the least RMSE as 0.4233. So, our improved model is lm(x110m ~ x100m+ long+ high).

### STEP3
### Drop one variable from the improved model from STEP2.
```{r}
set.seed(12345)
mod3_1 <- train(x110m ~ x100m+ long, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE3_1<-min(mod3_1$results$RMSE)  

set.seed(12345)
mod3_2 <- train(x110m ~ x100m+ high, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE3_2<-min(mod3_2$results$RMSE)  

set.seed(12345)
mod3_3 <- train(x110m ~long+ high, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE3_3<-min(mod3_3$results$RMSE)  

RMSE3_list<-c(RMSE3_1,RMSE3_2,RMSE3_3)
RMSE3_list
min(RMSE3_list)
```

*All of the model in step3 has larger RMSE than the RMSE of the final model in step2,(lm(x110m ~ x100m+ long+ high)). Therefore, our final model is lm(x110m ~ x100m+ long+ high), and RMSE is 0.4233.*

### 1500m

### full model
```{r}
set.seed(12345)
mod <- train(x1500m ~ x100m+ long+ shotput+ high+ x400m, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
min(mod$results$RMSE)
```

### STEP1 
### Drop one variable from the full model
```{r}
set.seed(12345)
mod1 <- train(x1500m ~ x100m+ long+ shotput+ high, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE1<-min(mod1$results$RMSE)  

set.seed(12345)
mod2 <- train(x1500m ~ x100m+ long+ shotput+ x400m, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE2<-min(mod2$results$RMSE)  

set.seed(12345)
mod3 <- train(x1500m ~ x100m+ long+ high+ x400m, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE3<-min(mod3$results$RMSE)  

set.seed(12345)
mod4 <- train(x1500m ~ x100m+ shotput+ high+ x400m, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))          
            
RMSE4<-min(mod4$results$RMSE)  

set.seed(12345)
mod5 <- train(x1500m ~ long+ shotput+ high+ x400m, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE5<-min(mod5$results$RMSE) 


RMSE_list<-c(RMSE1,RMSE2,RMSE3,RMSE4,RMSE5)
RMSE_list
min(RMSE_list)
```

*mod 2 has least RMSE 12.003. Now our improved model is lm(x1500m ~ x100m+ long+ shotput+ x400m)*


### STEP2
### Drop one variable from the improved model from STEP1.
```{r}
set.seed(12345)
mod2_1 <- train(x1500m ~ x100m+ long+ shotput, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE2_1<-min(mod2_1 $results$RMSE)  

set.seed(12345)
mod2_2 <- train(x1500m ~ x100m+ long+ x400m, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE2_2<-min(mod2_2$results$RMSE)  

set.seed(12345)
mod2_3 <- train(x1500m ~ x100m+ shotput+ x400m, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE2_3<-min(mod2_3$results$RMSE)  

set.seed(12345)
mod2_4 <- train(x1500m ~ long+ shotput+ x400m, 
             data=london,
             method="lm",
             trControl=trainControl(method="cv",number = 5),
             preProcess = c("center", "scale"))
RMSE2_4<-min(mod2_4$results$RMSE)  

RMSE2_list<-c(RMSE2_1,RMSE2_2,RMSE2_3,RMSE2_4)
RMSE2_list
min(RMSE2_list)
```

*All of the model in step2 has larger RMSE than the RMSE of the first model(lm(x1500m ~ x100m+ long+ shotput+ x400m)). Therefore, our final model is lm(x1500m ~ x100m+ long+ shotput+ x400m), and RMSE is 12.003*